The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-07-29 brains-introduction.tex
fix/todos: some minor rephrasings

\todo{add examples}

To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
authors spontaneously transform quotes, not only cropping them but also replacing words, when in fact they are implicitly required to copy them exactly.
We can therefore assume that most transformations, especially the simple ones, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.

>>>>

To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
authors spontaneously transform quotes, not only cropping them but also replacing words.
For instance the quote ``we will not be scared of these cowards'' (a substring of a quote from former Pakistani President Asif Ali Zardari) is also found as ``we will not be \textbf{afraid} of these cowards''.
More meaningful changes often happen too, such as the transformation of McCain's ``I admire Senator Obama and his accomplishments'' during the 2008 U.S. presidential campaign, into ``I \textbf{respect} Senator Obama and his accomplishments''.
Since authors are implicitly required to copy quotes exactly, we can assume that most transformations, especially simple ones such as those shown above, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.

### 2016-07-29 brains-protocol.tex
chore/all: typos, colors, todos
feat/discussion-conclusion: finish all the way to conclusion
fix/todos: some minor rephrasings

We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
\newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
%In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
%\todo{, as this would be akin to a drink from the firehose. \sf [Cam: supprimer? potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc. outside the scope etc. further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion? ou bien il s'agit d'une critique directe d'un reviewer ?  Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
Our goal, therefore, is to show that such effects exist and are measurable even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
%\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
As explained in the previous section, this is the reason we chose to use measures that can aggregate over all the transformations in the data set.}
\todo{this last sentence is unclear or not useful}

@@

We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
% utiliser "word-level measures", et/ou "For a given value of a given type of word-level measure (say, words of length 5), we define susceptibility as the ratio (...)"
\newtext{For a given group of words $g$ \todo{[not clear this is not a given word -- maybe present by example instead of generally]}, susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
That is:}\footnote{
\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
If, for instance, half the content words of a given quote are from group $g$, such a quote contributes .5 to the total $s_g^0$.}
}
$$\sigma_g = \frac{s_g}{s_g^0}$$

In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (e.g. for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.

@@

\todo{Relate to missed literature (\#15)}

@@

We \newtext{now examine} how words are modified \newtext{when} they are substituted, that is how their features change upon substitution.
Considering a word~$\wstart$ substituted for~$\warrival$, we measure how a feature~\newtext{$\phi$} of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we first average substitutions over each cluster, by considering the average of arrival word features for a given start word. \todo{explain this for susceptibility too}}
$$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$

@@

\todo{Relate to missed literature}

Results are gathered in Fig.~\ref{fig:feature-variations-global}.
A first observation is that all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, and the slope of $\nu_{\phi}$ is smaller than 1 in absolute value, independently of the feature considered.
\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the disappearing word, the appearing word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
\newtext{This reasoning is standard in the analysis of dynamical systems (where the same transformation is applied to the whole system over and over), and becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
the arrival value is always closer to the intersection with $y=x$, meaning that that intersection is an attractor point for the substitution process.
If the slope of $\nu_{\phi}$ were greater than one (in absolute value), the arrival value would always be \emph{farther} from the intersection than the start value was, making the intersection with $y=x$ a \emph{repulsor} point.
This is how the number of intersections with $y=x$ and the slope of $\nu_{\phi}$ at those intersections characterize the behavior of substitutions.
}}}
In other words, beyond individual variation patterns, the substitution process exhibits a unique attractor \newtext{for each feature}.

>>>>

We could therefore study the individual transformation process at work when authors alter quotations, by examining the modified words in each transformation.
\newtext{Since our approach is exploratory however, we do not know at the outset which precise effect of cognitive bias we are looking for.
Indeed the data we use does not come from a controlled experiment in the laboratory, designed to elicit a particular effect: they are recordings of real life interactions, with all the complexity and uncertainty of conditions this entails.
%In this study, therefore, we did not try to predict and explain in detail the cognitive processes responsible for transformations
%\todo{, as this would be akin to a drink from the firehose. \sf [Cam: supprimer? potentiellement un excès d'autoflagellation -- en fait on devrait pouvoir dire simplement: ``The prediction and detailed explanation of the cognitive processes responsible for transformations is outside of the scope of this study'', probablement après la fin de la phrase suivante (``of a larger complexity (the detailed prediction and deconstruction of the cognitive processes etc. outside the scope etc. further research)'' mais même là cette modestie pourrait plutôt être de mise dans la conclusion? ou bien il s'agit d'une critique directe d'un reviewer ?  Un peu pareil pour ``our approach is exploratory'' (que j'ai laissé au nom de l'ambiguïté d'exploratoire qui veut aussi bien dire ``premier brouillon'' que ``data mining'') et ``this first study'' (où j'ai enlevé le `first' car trop ``bon c'est à moitié fini mais on soumet quand même'' ;) ]}
Our goal, therefore, is to show that such effects exist and are measurable even if they are part of a larger complexity (the detailed prediction and deconstruction of the cognitive processes responsible for them being left to further research).
%\todo{That is, we aim to prove first a relatively weak hypothesis before trying stronger ones. [Cam: idem, je supprimerais celle-là]}
If this is confirmed, we will have successfully applied laboratory analyses to out-of-laboratory data, opening a path to explanations of actual (\hbox{vs.} simulated) cultural evolution with tools from cognitive science.
Aiming to exhibit such subtle biases in complex data is the main reason we chose to use aggregate measures that have a chance of smoothing out the possible variations of experimental conditions in the data set.}

@@

We say that a word is \emph{substitutable} if it appears in a quote which undergoes a substitution, whether the substitution operates on that word or on another one.
% utiliser "word-level measures", et/ou "For a given value of a given type of word-level measure (say, words of length 5), we define susceptibility as the ratio (...)"
\newtext{For a given group of words $g$, say all nouns, or all words in a particular range of values for a feature (e.g. words 2 to 4 letters long), susceptibility is computed as the ratio of $s_g$, the number of times words of that group are substituted, to $s_g^0$, the number of times words of that group would be substituted if substitutions fell randomly on substitutable words.
That is:}\footnote{
\newtext{$s_g^0$ is computed by summing, over all quotes undergoing a substitution, the ratio of the number of words in a quote that are from group $g$ to the number of words in the same quote that could have been the substitution's target.
If, for instance, half the content words of a given quote are nouns, such a quote contributes .5 to the total $s_{nouns}^0$.}
}
$$\sigma_g = \frac{s_g}{s_g^0}$$

In other words, susceptibility measures \newtext{how much more or less a group of words $g$ actually gets substituted compared to picking targets at random in quotes undergoing substitutions}.
\newtext{By applying this measure to Part-of-Speech (POS) categories and feature bins (e.g. for a feature $\phi$ and a bin $[a, b]$, $g = \{w | \phi(w) \in [a, b]\}$), susceptibility measures the bias in} the selection of start words involved in substitutions, \newtext{that is} the effect of word \newtext{properties} at the moment preceding the substitution when it is not yet known which word in the quotation will be substituted.\footnote{
\newtext{To avoid possible autocorrelation effects due to substitutions belonging to the same cluster (which are likely not statistically independent and may lead to overly optimistic confidence intervals), we scale $s_g$ and $s_g^0$ to count one for each cluster.
That is, each individual substitution contributes to $s_g$ or $s_g^0$ by only 1 over the number of substitutions in its cluster.}
}

@@

@@

We \newtext{now examine} how words are modified \newtext{when} they are substituted, that is how their features change upon substitution.
Considering a word~$\wstart$ substituted for~$\warrival$, we measure how a feature~\newtext{$\phi$} of~$\wstart$ varies when it is replaced with~$\warrival$, that is we look at~$\phi(\warrival)$ as a function of~$\phi(\wstart)$.
Averaging this value over all start words such that $\phi(w) = f$ yields the mean variation for that feature value~$f$, that is:\footnote{
\newtext{Similarly to what we do for susceptibility, we avoid possible autocorrelation effects by averaging start- and arrival-word features of substitutions from the same cluster into a single aggregate substitution per cluster.}
}
$$\nu_{\phi}(f) = \left<\phi(\warrival)\right>_{\left\lbrace \wstart\rightarrow\warrival | \phi(\wstart) = f \right\rbrace}$$

@@

Results are gathered in Fig.~\ref{fig:feature-variations-global}.
A first observation is that all graphs show the existence of a unique intersection of $\nu_{\phi}$ with $y=x$, and the slope of $\nu_{\phi}$ is smaller than 1 in absolute value, independently of the feature considered.
\newtext{This means that for each feature $\phi$, whichever the value $\phi(\wstart)$ of the disappearing word, the appearing word's feature value $\phi(\warrival)$ will, on average, be closer to that feature's intersection between $\nu_{\phi}$ and $y=x$.\footnote{
\newtext{This reasoning is standard in the analysis of dynamical systems (where the same transformation is applied to the whole system over and over), and becomes obvious when one manually simulates a substitution on the graph by picking a start value, using the $\nu_{\phi}$ curve to obtain the corresponding arrival value, and comparing it to the start value:
the arrival value is always closer to the intersection with $y=x$, meaning that that intersection is an attractor point for the substitution process.
If the slope of $\nu_{\phi}$ were greater than one (in absolute value), the arrival value would always be \emph{farther} from the intersection than the start value was, making the intersection with $y=x$ a \emph{repulsor} point.
This is how the number of intersections with $y=x$ and the slope of $\nu_{\phi}$ at those intersections characterize the behavior of substitutions.
}}}
In other words, beyond individual variation patterns, the substitution process exhibits a unique attractor \newtext{for each feature}.
\newtext{Note that this is also true under $\mathcal{H}_0$ or $\mathcal{H}_{00}$ (both null hypothesis curves have single intersections with $y=x$ with slopes smaller than 1):
the substitution process naturally leads to an attraction even under reasonable random conditions.}
